---
title: "My First Markdown Report - bioinformatics INR Final Project"
author: "Juan C. Sanchez-Arias"
date: "2021-07-11"
output:
  html_document: 
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
  pdf_document: 
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyr)
## setwd. NOTE: The working directory will be reset when the chunk is finished running.
# setwd("~/Documents/UofT/BCB330/Goldenberg_Lab/cbrworkshop/rmd_lesson/data")

## read in the same dataframe as we'd used earlier today
df2 = read.table("Example_data2_v2.txt", header = TRUE, as.is =TRUE)
```

# Investigating the dataframe
``` {r investigate, echo=FALSE}
names(df2)
head(df2)
dim(df2)
str(df2)
```

### Quick stats
* The `mean` age in our data is `r mean(df2$age)`
* The `std` of the age in our data is `r sd(df2$age)`
* The `range` of the ages in our data is [`r range(df2$age)`]

#### Create sex, site, and treated factor 
```{r new_factors, echo=FALSE}
df2$sex_factor = factor(df2$sex, levels = c("M","F"), labels = c("Male","Female"))
df2$site_factor = factor(df2$site, levels = c(1,2), labels = c("site 1","site 2"))
df2$treatment_factor = factor(df2$treated, levels = c(0,1), labels = c("control","treated"))
```


```{r aggregate_data, include=FALSE}
# Name grouping variable
aggregate(df2$age,list(df2$site),mean)
aggregate(df2$age,list(df2$site,df2$sex),mean)
aggregate(df2$age,
          list("site" = df2$site_factor,
               "sex" = df2$sex_factor), 
          mean)
```

### Exploring our new factors with tables
```{r tables, echo=FALSE}
knitr::kable(table(df2$treatment_factor, df2$sex_factor), 
             caption = 'A caption for our table.')
knitr::kable(table(df2$treatment_factor, df2$site_factor))
knitr::kable(table(df2$sex_factor, df2$site_factor))
# knitr::kable(df2$sex_factor)
```

#### Run a chi-square test to see if each category is randomly distributed relative to the others
```{r chisq, echo=FALSE}
chisq.test(table(df2$treatment_factor, df2$sex_factor))
chisq.test(table(df2$treatment_factor, df2$site_factor))
chisq.test(table(df2$sex_factor, df2$site_factor))
```

# Plotting our new factors
### Barplots of our new factors
```{r barplot1, fig.width=10, fig.height=4, echo=FALSE}
barplot(table(df2$treatment_factor, df2$sex_factor))
barplot(table(df2$treatment_factor, df2$site_factor))
```

``` {r barplot3, fig.width=10, fig.height=4, echo=FALSE}
barplot(table(df2$sex_factor, df2$site_factor))
```

### Legend-formatted barplot of our last table  
```{r barplot2, fig.width=10,fig.height=3, echo=FALSE}
## dress up one of the barplots 
barplot(table(df2$sex_factor, df2$site_factor),
        beside = TRUE,
        col = c("dodgerblue","darkorchid"),
        ylim = c(0,45),cex.names = 1.5,cex.axis = 1.5)
legend(x = 1, 45,
       legend = c("M","F"),
       fill = c("dodgerblue","darkorchid"), cex = 1.2, ncol=2)
```

# Plotting our marker data 
### Plotting a boxplot of each maker
```{r boxplot, fig.width=10,fig.height=4, echo=FALSE}
pairs(df2[,grep(pattern = "marker",x = names(df2))])
par(mfrow=c(1,5))

boxplot(marker1 ~ treatment_factor,data = df2)
boxplot(marker2 ~ treatment_factor,data = df2)
boxplot(marker3 ~ treatment_factor,data = df2)
boxplot(marker4 ~ treatment_factor,data = df2)
boxplot(marker5 ~ treatment_factor,data = df2)
```

### Reshaping our data to see all the markers on the same plot 
```{r reshape, echo=TRUE}
df2_mlt = tidyr::pivot_longer(data = df2, 
               cols = c("marker1","marker2","marker3","marker4","marker5"),
               names_to = "Marker")

```

## Our new boxplot 
```{r ggplot, fig.width=10,fig.height=4, fig.cap=paste("A caption for our figure"), echo=FALSE}
ggplot(df2_mlt,aes(x = treatment_factor, y = value)) +
  geom_boxplot(aes(fill = treatment_factor)) + 
  facet_grid(~Marker) + xlab("Treatment") + ylab("Marker (unit)") + 
  scale_fill_manual(values = c("control" = "turquoise","treated" = "orchid"),
                    name= "Treatment")+
  theme(axis.text.x=element_text(angle=45))

```

# T-tests on each marker
```{r t-tests, echo=FALSE}
t.test(marker1 ~ treatment_factor,data = df2)
t.test(marker2 ~ treatment_factor,data = df2)
t.test(marker3 ~ treatment_factor,data = df2)
t.test(marker4 ~ treatment_factor,data = df2)
t.test(marker5 ~ treatment_factor,data = df2)

```




 
  



